A better dataset?
I am using GSE143758, a mouse snRNA seq study with several samples, comparing
AD to WT mice. I am going to look at astrocytes and see if the data is high quality.

Loading in needed packages
```{r}
library(Seurat)
library(tidyverse)
library(readr)
```

Loading in the data into a csv so it can be turned into a seurat object
```{r}
data <- read.delim("/Users/yonatan/Desktop/single_cell_AD_WT_mice_data/GSE143758_Admouse_Crtx_7-10m_Astrocytes_UMIcounts.csv", sep = ';', row.names = 1)

```
turning the data into a seurat object
```{r}
ad_wt_astro <- CreateSeuratObject(data)
```
Adding in the percent of counts that map to mitochondrial DNA in the seurat object's meta data 
```{r}
ad_wt_astro[["percent.mt"]] <- PercentageFeatureSet(ad_wt_astro, pattern = "^MT-")
```
Generating key QC plots to help determine appropriate cutoff values for the gene count of cells, as well as the total number of couts for genes. These is to filter low quality cells, cells that are dying or low quality often have a lot of mitochondrial, and snRNA seq in theory should have no mitchondrial contamination.
```{r}
# Visualize QC metrics as a violin plot
VlnPlot(ad_wt_astro, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

