Loading in the needed libraries
```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(Seurat)
```
Loading in the actual data now
```{r}
AD_WT_DATA <- read.csv('/Users/yonatan/Desktop/Single_Cell_Proj/AQP4_snRNA_project/AQP4_Expression_snRNA_seq/Data/GSE138852_counts.csv', header=TRUE, row.names = 1)


```

Converting into a seurat object now, and loading in the metadata
```{r}
cell_metadata <- read.csv("/Users/yonatan/Desktop/Single_Cell_Proj/AQP4_snRNA_project/AQP4_Expression_snRNA_seq/Data/GSE138852_covariates.csv", header=TRUE, row.names = 1)
seurat_data <-  CreateSeuratObject(counts = AD_WT_DATA, meta.data = cell_metadata)
```
Preprocessing steps directly ripped from pbmc basic vignette, paramaters must be adjust for snRNA seq LATER!!!
```{r}
seurat_data[["percent.mt"]] <- PercentageFeatureSet(seurat_data, pattern = "^MT-")

head(seurat_data@meta.data, 5)
```

Basic Visulizations now
```{r}
# Visualize QC metrics as a violin plot
VlnPlot(seurat_data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.

plot1 <- FeatureScatter(seurat_data, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(seurat_data, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```
Based of visual inspection and in the interest of saving time (!! fix later) , i dont subset any nFeatureRNAm and I will use a 4 percent cutoff for mitcohrondial genes

!!! Edit, the paper already did QC and preprocessing for me, so I will just normalization and go from there. I might use Scttransform instead, since that approach might actually be better to use here.

Data normalization this was done the lazy way so look into fixing this later
```{r}
seurat_data <- NormalizeData(seurat_data, normalization.method = "LogNormalize", scale.factor = 10000)
```
After the data has been normalized I want to subset the seurat object to only contain Astrocyte cells

```{r}
#AD_CT_Astrocytes <- subset(seurat_data, subset="oupSample.celltype" == "astro" )

AD_CT_Astrocytes <- seurat_data[ , rownames(seurat_data@meta.data)[seurat_data@meta.data$oupSample.cellType == 'astro']]

```
Now that I have the normalized Astrocytes I can drop the batches that were considered faulty in the paper these should be the row names that end in AD7_AD8, and Ct7_Ct18
```{r}
# 1. Get the full list of cell barcodes (row names/column names)
cell_names <- colnames(AD_CT_Astrocytes)
# 2. Split each barcode by the underscore and keep the LAST element
# This separates the UMI from the Sample ID.
sample_ids <- sapply(
  X = strsplit(cell_names, split = "_"),
  FUN = function(x) {
    tail(x, 1) # Gets the last element after splitting
  }
)
clean_batch_astrocytes <- subset(AD_CT_Astrocytes, subset = '')
```

